Multilocus Genotype And Mycelial Compatability Group Assessment

### 2017-05-04 doc/RMD/mlg-mcg.Rmd
update mlgmcg

We can see that there are plenty of MLGs that contain multiple MCGs. This

```{r}
mlg.filter(dat11, dist = bruvo.dist, replen = other(dat11)$REPLEN) <- .Machine$double.eps + (3*(0.5/11))
dat11

fmcgmlg <- as.data.frame(table(mll(dat11, "contracted"), mll(dat11, "custom"))) %>%
  setNames(c("MLG", "MCG", "Freq")) %>%
  as_data_frame() %>%
  filter(Freq > 0)
fmcgs <- fmcgmlg %>%
  group_by(MCG) %>%
  summarize(MLGs = sum(Freq > 0), Samples = sum(Freq), Entropy = vegan::diversity(Freq), data = list(data_frame(MLG = MLG, Freq = Freq))) %>%
  arrange(desc(MLGs))
fmlgs <- fmcgmlg %>%
  group_by(MLG) %>%
  summarize(MCGs = sum(Freq > 0), Samples = sum(Freq), Entropy = vegan::diversity(Freq), data = list(data_frame(MCG = MCG, Freq = Freq))) %>%
  arrange(desc(Samples), desc(MCGs))
fmcgs
fmlgs
any(fmcgs$MLGs > 1)
```

@@

```{r}
make_mcgmlg_graph <- function(x){
  gdf <- mutate(x, MLG = paste0('MLG.', MLG))
  MLGS <- gdf %>% 
    group_by(MLG) %>%
    summarize(size = sum(Freq)) %>%
    rename(vertex = MLG)
  MCGS <- gdf %>% 
    group_by(MCG) %>%
    summarize(size = sum(Freq)) %>%
    rename(vertex = MCG)
  VAT <- bind_rows(MLGS, MCGS)
  g <- gdf %>% 
    select(MCG, MLG) %>%
    graph_from_data_frame(vertices = VAT)
  V(g)$type <- ifelse(grepl("MLG", V(g)$name), "Multilocus Genotype", "Mycelial Compatibility Group")
  g
}
g <- make_mcgmlg_graph(mcgmlg)
gf <- make_mcgmlg_graph(fmcgmlg)
osize <- V(g)$size
fosize <- V(gf)$size
```

@@

V(g)$size <- sqrt(osize)/10
V(gf)$size <- sqrt(fosize)/10
set.seed(2017-05-03)
lay2 <- create_layout(g, layout = "igraph", algorithm = "nicely")
flay2 <- create_layout(gf, layout = "igraph", algorithm = "nicely")

mcg_mlg_graph <- ggraph(lay2) +
  geom_node_circle(aes(r = size, fill = type)) +
  geom_node_text(aes(label = gsub("MLG.", "", name), color = type, size = size/4), show.legend = FALSE) +
  geom_edge_link(aes(start_cap = circle(node1.size, unit = "native"), 
                     end_cap = circle(node2.size, unit = "native")), 
                 arrow = arrow(length = unit(0.0125, "native"))) +
  coord_fixed() +
  scale_color_manual(values = c("white", "black")) +
  scale_fill_manual(values = c("black", "white")) +
  theme_graph(base_size = 16, base_family = "Helvetica") +
  theme(legend.position = "bottom") +
  ggtitle("Relation of Multilocus Genotypes and MCGs")

mcg_mlg_graph
mcg_mlg_graph %+% flay2
ggsave(file.path(PROJHOME, "results/figures/publication/FigureS2.svg"), 
       width = 88*3, height = 88*3, units = "mm")
```

@@

>>>>

@@

```{r}
make_mcgmlg_graph <- function(x){
  gdf <- mutate(x, MLG = paste0('MLG.', MLG))
  MLGS <- gdf %>% 
    group_by(MLG) %>%
    summarize(size = sum(Freq)) %>%
    rename(vertex = MLG)
  MCGS <- gdf %>% 
    group_by(MCG) %>%
    summarize(size = sum(Freq)) %>%
    rename(vertex = MCG)
  VAT <- bind_rows(MLGS, MCGS)
  g <- gdf %>% 
    select(MCG, MLG, Freq) %>%
    rename(weight = Freq) %>%
    graph_from_data_frame(vertices = VAT)
  V(g)$type <- ifelse(grepl("MLG", V(g)$name), "Multilocus Genotype", "Mycelial Compatibility Group")
  g
}
g <- make_mcgmlg_graph(mcgmlg)
osize <- V(g)$size
```

@@

V(g)$size <- sqrt(osize)/10
set.seed(2017-05-05)
lay2 <- create_layout(g, layout = "igraph", algorithm = "nicely")

mcg_mlg_graph <- ggraph(lay2) +
  geom_node_circle(aes(r = size, fill = type)) +
  geom_edge_link(aes(start_cap = circle(node1.size, unit = "native"), 
                     end_cap = circle(node2.size, unit = "native"), 
                     width = weight),
                 arrow = arrow(length = unit(0.01, "native"))) +
  geom_node_text(aes(label = gsub("MLG.", "", name), color = type, size = size/10), show.legend = FALSE) +
  coord_fixed() +
  scale_edge_width(range = c(0.25, 1.5)) +
  # scale_edge_color_viridis(option = "C", end = 0.9) +
  scale_color_manual(values = c("white", "black")) +
  scale_fill_manual(values = c("black", "white")) +
  theme_graph(base_size = 16, base_family = "Helvetica") +
  theme(legend.position = "bottom", legend.direction = "vertical") +
  ggtitle("Relation of Multilocus Genotypes and MCGs")

mcg_mlg_graph
ggsave(file.path(PROJHOME, "results/figures/publication/FigureS2.svg"), 
       width = 88*3, height = 88*3, units = "mm")
```

@@

There are a whole buttload of MLGs for those 5 MCGs. What are the severity 
ratings for those? Again, we take take a look at these AND simultaneously query
the top 5 MLGs for this.

```{r severity, fig.width = 7, fig.height = 4}

strat <- bind_cols(strata(dat11), 
                   other(dat11)$meta, 
                   data_frame(MLG = mll(dat11, "original")))
count_group <- . %>% 
  mutate(nobs = n()) %>%  # count the number of samples/MCG
  ungroup() %>%
  arrange(desc(nobs)) # arrange by number of samples and reorder factors 
Severity <- filter(strat, MCG %in% mcgs$MCG[1:5]) %>%
  group_by(MCG) %>% 
  count_group %>%
  mutate(MCG = forcats::fct_inorder(factor(MCG), ordered = TRUE)) %>%
  mutate(MLG = ifelse(MLG %in% mlgs$MLG[1:5], paste("MLG", MLG), "Other"))

severity_plot <- ggplot(Severity, aes(x = MCG, y = Severity)) +
  geom_point(aes(fill = MLG), 
             position = position_jitter(width = 0.2),
             alpha = 0.75,
             pch = 21) +
  scale_fill_viridis(discrete = TRUE, direction = -1) +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(aspect.ratio = 0.6) +
  facet_wrap(~Region, nrow = 2) +
  ylim(c(3.5, 8)) +
  labs(list(
    title = "Severity by MCG and Region",
    fill = "Multilocus Genotype",
    subtitle = "Five most abundant multilocus genotypes shown"
  ))
severity_plot
ggsave(severity_plot, filename = "results/figures/publication/FigureS3.svg", width = 183, unit = "mm")
```
